---
title: "Hand R DataVis Final Webpage"
author: "Andrew Hand"
date: "01/05/2025"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data Visualization in R: Hand Directory {.tabset .tabset-fade .tabset-pills}

## Data Vis Modules {.tabset .tabset-fade .tabset-pills}
#### This section includes code from all past modules from Data Visualization in R course. 

### Module 1
#### GGPLOT Basics

```{r,message=F, warning=FALSE}
# List any packages you need to use here
packages <- c("ggplot2", "readr", "tidyverse", "dplyr", "ggpubr")

#Check to see if any of your listed packages need installed
check_install_packages <- function(pkg){
  if (!require(pkg, character.only = TRUE)) {
    install.packages(pkg, dependencies = TRUE)
    library(pkg, character.only = TRUE)
  }
}

# Download the packages and read in the libraries if necessary
sapply(packages, check_install_packages)

data("USArrests")
view(USArrests)
head(USArrests)
?USArrests
```

USArrests data questions: 
1: What are the variables available? 
murder, assault, urbanpop, and rape 

2: How is each variable defined or calculated? 
muder, assault, and rape are numeric values calculated for one per 100,000 arrests
urbanpop is defined as the percent urban population

3: Is each one numerical or categorical? 
each is numerical 

Horsepower Plot:
```{r,message=F, warning=FALSE}
# libraries 
library(ggplot2)
#General format is going to be calling a ggplot, followed by the dataframe name (mtcars), followed by defining the X and Y variables of the graphic.
data("mtcars")  
ggplot(mtcars, aes(x = mpg, y=hp)) +
    #You then indicate the type of graph to make (in this case, a dotplot using points).
    geom_point() +
  #changing dots to size 2.4 and star shaped
  geom_point(size = 2.4, shape = 8) +
  #change theme to minimal 
  theme_minimal() +
  #create cyl column into a color gradient 
  aes(colour = cyl) +
  #change the legend position
  theme(legend.position = "bottom") +
  #labeling the graph
  labs(title = "Effect of Horsepower on Fuel Efficiency", subtitle = "Categorized by Number of Cylinders", x = "Horsepower", y = "Fuel Efficiency (MPG)")
```

Iris Plot:
```{r,message=F, warning=FALSE}
#using a different demo data set to create another graph 
library(ggplot2)
data(Irisi)
#create plot
ggplot(iris, aes(x = Species, y = Petal.Length, fill = Species)) + geom_boxplot() +
  #change theme
  theme_classic() +
  #labels
  labs(title = "Patel Length of Iris by Species") + 
  #custom legend position
  theme(legend.position = c(0.2,0.8))
```

### Module 2 
#### Scatter Plots, Dot Plots, Strip Charts, and Line Plots

task 1: Scatter Plot and Regression Line: 
```{r,message=F, warning=FALSE}
#Load data 
data("USArrests")
#view the data set 
head(USArrests)
#load proper packages
library(ggplot2)
#create a scatter plot with the chosen x and y variables
ggplot(USArrests, aes(y = Assault, x =Murder)) + geom_point() +
  #set theme 
  theme_classic() + 
  #add labels 
  labs(title = "Scatter Plot of Assault vs. Murder Rates", x = "Murder Rate", y = 
         "Assault Rate") + 
  #add a linear model regression line 
    geom_smooth(method = lm, color = "black", fill = "red") + 
  #changing aesthetics 
  geom_point(size = 3.5, shape = 21, fill = "lightblue") + 
  #change the font sizes a little 
  theme(plot.title = element_text(size=16, face = "bold"), axis.title.x = element_text(size = 14), axis.title.y = element_text(size = 14))
```

Task 2: Challenge Line Plot Creation:
```{r,message=F, warning=FALSE}
#create a new column for the states
USArrests$State <- rownames(USArrests)
#creating a column for the means
USArrests$AverageCrimeRate <- rowMeans(USArrests[, c("Murder", "Assault", "Rape")])

# Create the line plot
ggplot(data = USArrests, aes(x = State, y = AverageCrimeRate, group = 1)) +
  geom_line(color = "steelblue", size = 1) +  # Line
  geom_point(color = "red", size = 3) +   # Points
  labs(title = "Line Plot of Average Crime Rate by State",
       x = "State",
       y = "Average Crime Rate") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))  # Rotate x-axis labels

#notes: couldn't figure out how to add the states to the x axis the first time
``` 

### Module 3
#### Boxplots and Violon Plots

```{r problem_fix, message=F, warning=FALSE, include=FALSE}
library(readr)
library(tidyverse)

CAM <- read_csv("Violin_Plot_Data.csv")
data_long <- CAM %>%
  pivot_longer(
    cols = starts_with("Repeat"),
    names_to = "Repeat",
    values_to = "values"
  )
min_value <- min(data_long$values, na.rm = TRUE)
max_value <- max(data_long$values, na.rm = TRUE)
```

Replica With Half Violin Plot:
```{r replica_half_violin, message=F, warning=FALSE}
library(readr)
CAM <- read_csv("Violin_Plot_Data.csv")
View(CAM)
#load all packages needed for this module
library("ggplot2") 
library("readr") 
library("tidyverse")
library("dplyr") 
library("ggpubr") 
library("see")
library("scales")

#view the data
head(CAM)
#Calculate min and max
min_value <- min(data_long$values, na.rm = TRUE) 
max_value <- max(data_long$values, na.rm = TRUE)

#give your newly formatted data a name you will recognize, in this case "data_long"
data_long <- CAM %>%
  #Pivot the data from having many columns to many rows
  pivot_longer(
    cols = starts_with("Repeat"),  # Select columns to pivot
    names_to = "Repeat", 
    values_to = "values") #give the newly created column a name
#view the resulting data
head(CAM) 
#calculating min and max 
 min_value <- min(data_long$values, na.rm = TRUE) 
max_value <- max(data_long$values, na.rm = TRUE) 
#writing ggplot2 base data
ggplot(data_long, aes(x = F1Performance, y = values)) + geom_jitter(size = 6, width = 0.1, height = 0, alpha = NA, aes(color = F1Performance)) + geom_violinhalf() + 

  #add in the color and legend 
  aes(fill = F1Performance) +
  #making the color transparent
  aes(alpha = "0.20", size.f = 2) +
  #adding quantile lines
  geom_violinhalf(draw_quantiles = c(0.25, 0.5, 0.75)) + 
  #flip the axis 
  coord_flip() + 
  
  #changing the color to orange and purple
  scale_fill_manual(values = c("purple4", "darkorange2")) + 
  scale_color_manual(values = c("purple4", "darkorange2")) +
  #add summary statistic and highlight it
  stat_summary(fun = median, geom = "point", shape = 21, size = 3, fill = "white", color = "black", stroke = 1.5, alpha = NA) + 
  #changing theme 
  theme_minimal() +
  #changing y axis
  theme(axis.title.y = element_blank(), axis.text.y = element_blank(), axis.ticks.y = element_blank(), axis.line.y = element_blank()) +
  #changing x axis
  theme(axis.line.x = element_line(color = "black", size = 2)) + 
  #changing major and minor grid lines 
    # Remove major grid lines for y axis
  theme(panel.grid.major.y = element_blank(),
      # Remove minor grid lines for x axis
  panel.grid.minor.x = element_blank(),     
    #changing the major x axis grid lines
  panel.grid.major.x = element_line(color = "grey", linetype = "dashed", size = 1.5)) +
  #adding title
  theme(plot.title = element_text(hjust = 0.5, face="bold")) + 
    theme(legend.position = "none") +
  
  #adding text labels 
  geom_text(aes(x = "SVMWithGradCAMMaps", label = "SVM + GRAD-CAM++", y = 0.64), vjust = -3.5, color = "darkorange2", size = 4.5) + 
  geom_text(aes(x = "SMVWithDeepShapMaps", label = "SMV + Deep SHAP", y = 0.6), vjust = -14, color = "purple4", size = 4.5) +
  scale_y_continuous(limits = c(min_value, max_value), breaks = seq(min_value, max_value, by = 0.02)) +
  #adding a title and changing axis name 
  labs(title = "Fig. 7. Grad-CAM++ saliency maps capture unique predictive information.", face = "bold") +
  labs(y = "F1")
  
#that was really difficult
    
```

Replica With Full Violin Plot:
```{r full_violin, message=F, warning=FALSE}
library(readr)
CAM <- read_csv("Violin_Plot_Data.csv")
View(CAM)

#load all packages needed for this module
library("ggplot2") 
library("readr") 
library("tidyverse")
library("dplyr") 
library("ggpubr") 
library("see")
library("scales")

#view the data
head(CAM)
#Calculate min and max
min_value <- min(data_long$values, na.rm = TRUE) 
max_value <- max(data_long$values, na.rm = TRUE)

#give your newly formatted data a name you will recognize, in this case "data_long"
data_long <- CAM %>%
  #Pivot the data from having many columns to many rows
  pivot_longer(
    cols = starts_with("Repeat"),  # Select columns to pivot
    names_to = "Repeat", 
    values_to = "values") #give the newly created column a name
#view the resulting data
head(CAM) 
#calculating min and max 
 min_value <- min(data_long$values, na.rm = TRUE) 
max_value <- max(data_long$values, na.rm = TRUE) 
#writing ggplot2 base data
ggplot(data_long, aes(x = F1Performance, y = values)) + geom_jitter(size = 6, width = 0.1, height = 0, alpha = NA, aes(color = F1Performance)) + geom_violin() + 

  #add in the color and legend 
  aes(fill = F1Performance) +
  #making the color transparent
  aes(alpha = "0.20", size.f = 2) +
  #adding quantile lines
  geom_violin(draw_quantiles = c(0.25, 0.5, 0.75)) + 
  #flip the axis 
  coord_flip() + 
  
  #changing the color to orange and purple
  scale_fill_manual(values = c("purple4", "darkorange2")) + 
  scale_color_manual(values = c("purple4", "darkorange2")) +
  #add summary statistic and highlight it
  stat_summary(fun = median, geom = "point", shape = 21, size = 3, fill = "white", color = "black", stroke = 1.5, alpha = NA) + 
  #changing theme 
  theme_minimal() +
  #changing y axis
  theme(axis.title.y = element_blank(), axis.text.y = element_blank(), axis.ticks.y = element_blank(), axis.line.y = element_blank()) +
  #changing x axis
  theme(axis.line.x = element_line(color = "black", size = 2)) + 
  #changing major and minor grid lines 
    # Remove major grid lines for y axis
  theme(panel.grid.major.y = element_blank(),
      # Remove minor grid lines for x axis
  panel.grid.minor.x = element_blank(),     
    #changing the major x axis grid lines
  panel.grid.major.x = element_line(color = "grey", linetype = "dashed", size = 1.5)) +
  #adding title
  theme(plot.title = element_text(hjust = 0.5, face="bold")) + 
    theme(legend.position = "none") +
  
  #adding text labels 
  geom_text(aes(x = "SVMWithGradCAMMaps", label = "SVM + GRAD-CAM++", y = 0.64), vjust = -3.5, color = "darkorange2", size = 4.5) + 
  geom_text(aes(x = "SMVWithDeepShapMaps", label = "SMV + Deep SHAP", y = 0.6), vjust = -14, color = "purple4", size = 4.5) +
  scale_y_continuous(limits = c(min_value, max_value), breaks = seq(min_value, max_value, by = 0.02)) +
  #adding a title and changing axis name 
  labs(title = "Fig. 7. Grad-CAM++ saliency maps capture unique predictive information.", face = "bold") +
  labs(y = "F1")
  
```

Module 3 Challenge Violin Plus Box Plot: 
```{r challenge_plot, message=F, warning=FALSE}
# libraries
library("rmarkdown")
library("knitr")

# libraries 
library("ggplot2") 
library("readr") 
library("tidyverse")
library("dplyr") 
library("ggpubr") 
library("see")
library("scales") 
library("hrbrthemes") 
library(readr)

CAM <- read_csv("Violin_Plot_Data.csv")
View(CAM)

# calculate min and max values
min_value <- min(data_long$values, na.rm = TRUE) 
max_value <- max(data_long$values, na.rm = TRUE)

# make the data long format 
data_long <- CAM %>%
  #Pivot the data from having many columns to many rows
  pivot_longer(
    cols = starts_with("Repeat"),  # Select columns to pivot
    names_to = "Repeat", 
    values_to = "values") #give the newly created column a name
head(data_long)

# create the base ggplot
ggplot(data_long, aes(x = F1Performance, y = values)) + 
  # add data points 
  geom_jitter(size = 5, width = 0.1, height = 0, alpha = NA) +
  # add violin and boxplot 
  geom_violin(aes(fill = F1Performance, alpha = 0.2)) +
  geom_boxplot(width = 0.3, aes(color_fill_manual = "grey", alpha = 0.2)) + 
  # add a mean point 
  stat_summary(fun = median, geom = "point", shape = 21, size = 3, fill = "white", color = "black", stroke = 1.5, alpha = NA) + 
  # changing theme 
  theme_classic() + 
  theme(legend.position = "none") + 
  theme(axis.title.y = element_blank()) + 
  labs(title = "Fig. 7. Grad-CAM++ saliency maps capture unique predictive information.") 
  
  
# doing this a second time was a little easier 

```

### Module 4 
#### Barplots and Error Bars

mtcars Barplot with Error Bars:
```{r barplots_and_errorbars, message=F, warning=FALSE}

# libraries
library(ggplot2)
library(dplyr)
mean_mpg <- mean(mtcars$mpg)

# Task 1
# load data set mtcars 
data(mtcars)
head(mtcars)
group_data <- mtcars %>% group_by(cyl)
summary(mtcars)
# calculating mean 
mean_mpg <- mean(mtcars$mpg) + 
  print(mean_mpg)
# mean_mpg = 20.0962
# calculating Standard Error from SD 
sd_mpg <- sd(mtcars$mpg, na.rm = TRUE) 
# creating a sample size vriable 
n <- sum(!is.na(mtcars$mpg))
se_mpg <- sd_mpg / sqrt(n)
print(se_mpg)
# SE_mpg = 1.065
# summarize data in new dataframe 
data_summary <- mtcars %>% group_by(cyl) %>% summarise(se_mpg = sd_mpg / sqrt(n), mean_mpg = mean(mpg))
print(data_summary)

# Task 2 creating a bar plot 
ggplot(data_summary, aes(x = factor(cyl), y = mean_mpg,)) + geom_col(fill = c("skyblue", "orange", "green"), width = 0.7) +
  geom_errorbar(aes(ymin = mean_mpg - se_mpg, ymax = mean_mpg + se_mpg), width = 0.25, size = 2) + 
  theme_classic() + 
  labs(title = "Bar Plot of Mean MPG by Cylinders", x = "Number of Cylinders", y = "Mean MPG")
  
#messed up on the summary stats at first but fixed it 

```  

### Module 5 
#### Density Plots, Histograms and Frequency Polygons
 
2D Density Graphics Plot:
```{r chunk 1, message=F, warning=FALSE}
# load packages 
library(readr)
library(ggplot2)
library(dplyr)
# read in csv and rename it via import dataset function
library(readr)
population_data <- read_csv("log_population_data.csv") 
View(population_data) 
head(population_data)
# create the density plot
ggplot(population_data, aes(x = Log10_Current_Population, y = Log10_Past_Population)) +
  stat_density_2d(aes(fill = after_stat(level)), geom = "polygon", color = "white") +
  scale_fill_gradient(low = "darkblue", high = "skyblue", name = "level") + 
  # I couldn't figure out how to use the scale_fill_distiller function so I used a differnt one
  
  theme_minimal() + 
  labs(title = "2D Density Plot of Population Sizes", x = "Log10(Current population size N0", y = "Log10(Past population ize N1)") 
```

Adding Density Plot To The Margins of A Different Plot:
```{r chunk 2, message=FALSE, warning=FALSE}
library(readr)
library(ggplot2)
library(dplyr)
library(ggExtra)
library(readr)
longevity_data <- read_csv("longevity_data.csv")
View(longevity_data)
head(longevity_data)
# put the data into log format
long <- longevity_data %>% #create a new dataframe called "long" that contains all your newly calculated variables
  mutate( #mutate tells the program to perform new calculations
    log_mass = log10(mass_g),                          # create a new column called "log_mass" which Log-transforms mass values
    log_lifespan = log10(maximum_lifespan_yr))  %>%          # create a new colummn called "log_lifespan" that Log-transforms lifespan value
   group_by(order) %>%        # this tells it that after "mutate", you are going to start a new function. for each "order" or group of animals    
  mutate(order_size = n())      #calculate the sample size of each order and put it in a column called "order_size". 

#Now you have a sample size for each order, and you have transformed each mass and lifespan value to log form. 

head(long)
# create a dotplot 
p = ggplot(long, aes(x =log_mass, y = log_lifespan, color = class, size = order_size)) + 
  #makign the points transparent by 30%
  geom_point(alpha = 0.3) +
  #regression line
  geom_smooth(method = "lm", se = FALSE, linetype = "solid", aes(color = class)) +
  # change the color scheme
  scale_color_manual(values = c("lightgreen", "darkslategray")) + 
  # change labels and theme
  labs(title = "Bubble Chart of Longevity and Body Mass", x = "Log (Body Mass [g])", y = "Log (Maximum Lifespan [yr])") + 
  theme_minimal() + 
  # remove legends 
  theme(legend.position = "none") +
  # changing axis and title labels theme 
  theme(plot.title = element_text(size = 12, face = "bold"), 
        axis.title.x = element_text(size = 12, face = "bold"), 
        axis.title.y = element_text(size = 12, face = "bold")) + 
  # add text annotations inside the graph
  annotate("text", label = "Aves", x = 5.6, y = 1.9, color = "lightgreen", size = 5, face = "bold") + 
  annotate("text", label = "Mammals", x = 6.5, y = 1.4, color = "darkslategray", size = 5, face = "bold")

ggExtra::ggMarginal(p, type = "density", groupFill = TRUE, alpha = 0.4)

```

Interpretation Questions: 
1) adding the density plots in the margins adds another visualization tool that
shows how the different classes are distributed in the x and y axis. 

2) 
  1: log mass plotted on the x axis.
  2: log lifespan plotted on the y axis.
  3: making the classes differnet colors.
  4: order size depicts the sample size of the populations.
  5: transparency adds a visual aid allowing you to see more dense populations.
  6: marginal density plots add another visual to detemrine where populations 
  are the most dense.
  6 1/2: regression lines show the trend of the two classes.

3) the more body mass = longer lifespan. More extreme in Aves as seen by the 
trendlines.

4) for Aves, there are more samples in the lower left of the graph (low body mass and short lifespan), compared to other areas of the graph. This could be a possible bias since most of the data collected was of that low body mass and short lifespan category. The same can be sais for the Mammals where the data is more accumulated towards a high lifespan and body mass. 

5) We could use the dodge function to create 2 side-by-side graphs that might make the graph easier to read. Or, we could change "se = TRUE" to show error along the trend lines. 

Create Your Own Plot Challenge:
```{r chunk 3, message=FALSE, warning=FALSE}
# load packages 
library(ggplot2)
library(dplyr)
library(ggExtra)
# using CO2 dataset 
data("CO2")
head(CO2)

p1 = ggplot(CO2, aes(x = conc, y = uptake, color = Plant, fill = Treatment, linetype = Treatment)) +
  # Line plot for trends
  geom_point(aes(group = interaction(Plant, Treatment)), size = 1, alpha = 0.8) + geom_line() +
  # set theme 
  theme_minimal() + theme(legend.position = "bottom") +
  # set labels 
  labs(title = "CO2 Uptake Across Plant Origin, Treatment, and Concentration", x = "Concentration of CO2(mL/L)", y = "Uptake of CO2 (μmol/m^2 sec)") + 
  theme(plot.title = element_text(face = "bold"), 
        plot.title.x = element_text(face = "bold"), 
        plot.title.y = element_text(face = "bold"))  
  ggMarginal(p1, type = "density", groupFill = TRUE, alpha = 0.4) 
``` 

### Module 6 
#### Integrating Multiple Charts
 
Module 6 Notes:
export a figure as a pdf, eps, or png: 
ggexport(figure, filename, "figure1.pdf", nrow = 2, ncol = 1) 

to save a plot in r: 
save_plot("plot name.pdf", plot name, ncol = x, nrow = 2, base_aspect_ratio = 1.4) # always 1.4 to add room for legend

Chick Weight Plot Code:
```{r chick_plot_code, message=FALSE, warning = FALSE}
# load libraries
library(ggplot2)   # facet_grid(), facet_wrap() for creating multiple figures with the same axis  
library(cowplot)   # plot_grid() and draw_plot(plot, x = , y = , width = , height = )
library(gridExtra) # grid.arrange()
library(ggpubr)    # ggarrange() easiest to use
library(patchwork) # plot_layout()
library(dplyr)



# load dataset 

data(ChickWeight)
head(ChickWeight)

# summary statistics
summary(ChickWeight)


# create plot   
p <- ggplot(ChickWeight, aes(x = Time, y = weight, color = Chick, alpha = 0.1)) + geom_line() +

# regressChickWeight# regression line 
geom_smooth(color = "black", size = 1.2) +
  
# labels 
labs(title = "Chick Growth by Diet Type", x = "Time (Days)", y = "Weight (Grams)") +
  
# theme and legend
theme_minimal() + theme(legend.position = "none") +
  
# facet wrap 
facet_wrap(vars(Diet), ncol = 4, nrow = 1) 

```

Output of Chick Plot:
```{r plot_p, message=FALSE, warning=FALSE}
plot(p)
``` 

Multipanel Figure Plot:
```{r multipanel_plot, message=FALSE, warning = FALSE}
# load packages 
library(dplyr)
library(ggplot2)   # facet_grid(), facet_wrap() for creating multiple figures with the same axis  
library(cowplot)   # plot_grid() and draw_plot(plot, x = , y = , width = , height = )
library(gridExtra) # grid.arrange()
library(ggpubr)    # ggarrange() easiest to use
library(patchwork) # plot_layout()
# load data befor each plot  
data(CO2)
head(CO2)

# 1. violin plot
v <- ggplot(CO2, aes(x = Treatment, y = uptake, alpha = 0.4, fill = Type)) +
  geom_jitter(size = 3, alpha = NA, position = position_jitterdodge(jitter.width = 0.2, dodge.width = 1)) +
  geom_violin(position = position_dodge(1), aes(alpha = 0.2)) + 
  geom_boxplot(width = 0.3, position = position_dodge(1), aes(color_fill_manual = "grey", alpha = 0.2)) + 
# i don't know why i can't change the fill of my boxplots to be grey

  # labels 
labs(title = "CO2 Uptake of Different Plant Type by Treatment", x = "Treatment", y = "CO2 Uptake (μmol/m^2 sec)") + 
  
  # theme # theme Type
theme_classic() + theme(plot.title = element_text(face = "bold"), 
                        axis.title.x = element_text(face = "bold"), 
                        axis.title.y = element_text(face = "bold")) +
  
# remove alpha legend
guides(alpha = "none")
v



# 2. Line plot 
l <- ggplot(CO2, aes(x = conc, y = uptake)) + geom_line(aes(color = Treatment, fill = Treatment, size = 2, alpha = 0.8)) +
  
  # facet 
facet_grid(cols = vars(Type)) +
  
# labels 
labs(title = "Ambient CO2 Uptake in Different Areas and CO2 Concentration by Treatment", x = "CO2 Concentration (mL/L)", y = "CO2 Uptake (μmol/m^2 sec)") +

# themes 
theme_minimal() + 
  theme(plot.title = element_text(face = "bold", hjust = 0.5), 
                        axis.title = element_text(face = "bold")) +

# remove alpha and size legend
guides(alpha = "none", size = "none")
l



# 3. density plot with a vertical dashed line at mean

# create the plot 
d <- ggplot(CO2, aes(x = uptake, fill = Treatment)) +
  geom_density(alpha = 0.5) +  
  
# add a vertical line at mean 
geom_vline(aes(xintercept = mean(uptake, na.rm = TRUE)), 
             linetype = "dashed", color = "black", size = 1) +  
  
# labels 
labs(title = "Density Plot of CO2 Uptake by Treatment with Mean Line", x = "CO2 Uptake", y = "Density") +

# theme 
  theme_minimal() + theme(plot.title = element_text(face = "bold"), 
                          title.axis = element_text(face = "bold"))

d


# combine plots into one figure 

multipanel_plot1 <- plot_grid(
  v + theme(plot.title = element_blank(), axis.title.x = element_blank(), axis.title.y = element_blank()),
  l + theme(plot.title = element_blank(), axis.title = element_blank()),
  d + theme(plot.title = element_blank(), axis.title = element_blank()),
  ncol = 1,                        # Arrange in 1 column
  labels = c("A", "B", "C"),       # Add labels to each plot
  label_size = 13, 
  label_y = c(1.03, 1.05, 1.12), 
  rel_heights = c(1.2, 1, 0.9))

# print multipanel_plot1
multipanel_plot1
```

## Final Project Part 2 {.tabset .tabset-fade}
#### This section includes 5 tabs, each includong data visualization of 3 datasets giving different insights into current COVID-19 case data in Maryland and St. Mary's County. Links to the Maryland.gov webpage with the specific datasets used in each data visualization are provided at the start of each tab. 

### Maryland COVID-19 Cases by County
#### Maryland COVID-19 Cases by County

Link to data: https://opendata.maryland.gov/d/tm86-dujs
```{r County, message=F, warning=F}
# libraries 
library(ggplot2)
library(tidyr)
library(dplyr)
library(readr)
library(tidyverse)
library(ggExtra)

# load dataset
data <- read.csv("mdcovid_cases_by_county.csv")

# convert DATE to proper format
data$DATE <- as.Date(data$DATE, format = "%m/%d/%Y")

# select relevant columns
selected_data <- data[, c("DATE", "St_Marys", "Calvert", "Charles")]

# convert data to long format
long_data <- pivot_longer(selected_data, 
                          cols = c(St_Marys, Calvert, Charles), 
                          names_to = "County", 
                          values_to = "Cases")

# create the plot
countyp <- ggplot(long_data, aes(x = DATE, y = Cases)) +
  geom_point(aes(color = County)) +

# labels
  labs(title = "COVID-19 Cases in St. Mary's, Calvert, and Charles Counties",
       x = "Date",
       y = "Number of Cases",
       color = "County") +
# theme 
  theme_minimal() + theme(plot.title = element_text(face = "bold"))  + theme(legend.position = c(0.1,0.81), 
                                                                             legend.title = element_text(size = 13), 
                                                                             legend.background = element_rect(fill = "white", color = "black"))
  
# add margin density plot 
  ggMarginal(countyp, type = "density", margins = "y", groupColour = TRUE, groupFill = TRUE)

```

### Maryland COVID-19 Cases by Age Group
#### Maryland COVID-19 Cases by Age Group

Link to data: https://opendata.maryland.gov/d/sjqg-bqsu
```{r Age, message=F, warning=F}
# libraries 
library(ggplot2)
library(tidyr)
library(dplyr)
library(readr)
library(tidyverse)
library(ggExtra)
# read in the data 
data <- read.csv("mdcovid_cases_by_age.csv")

# remove the "Age_Unknown" column
age <- data[, !names(data) %in% c("Age_Unknown")]

# gather the data into a long format for ggplot2
age <- data %>%
  select(-c(OBJECTID, DATE)) %>%
  gather(key = "Age_Group", value = "Cases")

# create the histogram plot
agep <- ggplot(age, aes(x = Age_Group, y = Cases, fill = Age_Group)) +
  geom_bar(stat = "identity") +

  # add labels 
  labs(title = "COVID-19 Cases by Age Group", x = "Age Group", y = "Number of Cases") +

  # change theme
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "none")

# print plot 
agep

```

### Maryland COVID-19 Cases by Gender
#### Maryland COVID-19 Cases by Gender

Link to data: https://opendata.maryland.gov/d/py3p-2bgq
```{r Gender, message=F, warning=F}
# libraries 
library(ggplot2)
library(tidyr)
library(dplyr)
library(readr)
library(tidyverse)
library(ggExtra)

# read in the dataset
data <- read.csv("mdcovid_cases_by_gender.csv")

# remove the "Unknown" column because it has no data
data <- data[, !names(data) %in% c("Unknown")]

# convert the data into a long format
long_data <- data %>%
  select(-c(OBJECTID, DATE)) %>%
  pivot_longer(cols = everything(), names_to = "Gender", values_to = "Cases")

# make sure data is in long format 
print(head(long_data))

# create the violin plot combined with boxplot and jitter
pgender <- ggplot(long_data, aes(x = Gender, y = Cases, fill = Gender)) +
  geom_jitter(size = 2, width = 0.1, alpha = 0.9) +
  geom_violin(trim = FALSE, alpha = 0.6) +
  geom_boxplot(width = 0.2, outlier.shape = NA, alpha = 0.6) +
  
# labels
  labs(title = "COVID-19 Cases by Gender", x = "Gender", y = "Number of Cases") +

# theme
  theme_minimal() +
  theme(legend.position = "none", plot.title = element_text(face = "bold"))

# Print the gender plot
pgender


```

### Maryland COVID-19 Cases by Ethnicity
#### Maryland COVID-19 Cases by Ethnicity

Link to data: https://opendata.maryland.gov/d/xnfm-sgpt 
```{r Ethnicity, message=F, warning=F}
# libraries 
library(ggplot2)
library(tidyr)
library(dplyr)
library(readr)
library(tidyverse)
library(ggExtra)

# read in the dataset
data <- read.csv("mdcovid_cases_by_ethnicity.csv")

# select the most recent data (last row of the dataset)
latest_data <- tail(data, 1)

# remove out of date columns
ethnicity_data <- latest_data[, c("African_American", "White", "Hispanic", "Asian", "Other")]

# calculate the percentage for each ethnicity
ethnicity_percent <- colSums(ethnicity_data) / sum(ethnicity_data) * 100

# prepare data for plotting
ethnicity_df <- data.frame(
  Ethnicity = names(ethnicity_percent),
  Percentage = ethnicity_percent
)

# create the pie chart
ethp <- ggplot(ethnicity_df, aes(x = "", y = Percentage, fill = Ethnicity)) +
  geom_bar(stat = "identity", width = 1) +

# this actually makes the bar graph into a pie chart
  coord_polar("y", start = 0) +

# adds the percentages 
  geom_text(aes(label = sprintf("%.1f%%", Percentage)), 
            position = position_stack(vjust = 0.5), 
            size = 4) +

# labels
  labs(title = "COVID-19 Cases by Ethnicity") +

# theme
  theme_void() +
  theme(legend.position = "right", plot.title = element_text(face = "bold")) +
  scale_fill_brewer(palette = "Set3")

# print plot 
ethp

```

### Multipanel Plot of Maryland COVID-19 Cases by Age Group, Gender, and Ethnicity
#### This is a multipanel plot of maryland COVID-19 cases by age group, gender, and ethnicity. Seperate plots can be found in the previouse tabs along with the link from which the data was aquired.  

```{r Multipanel, message=F, warning=F}
# libraries 
library(ggplot2)
library(tidyr)
library(cowplot)


multipanel_plot <- plot_grid(
  agep + theme(plot.title = element_blank(), axis.title.x = element_blank(), axis.title.y = element_blank()),
  ethp + theme(plot.title = element_blank(), axis.title = element_blank()),
  pgender + theme(plot.title = element_blank(), axis.title = element_blank()),
  ncol = 2,                        # Arrange 2 column
  labels = c("A", "C", "B"),       # Add labels to each plot
  label_size = 12, 
  label_y = c(1, 1, 1.05),
  rel_heights = c(1, 1, 1))

#print multipanel plot
multipanel_plot

```
